---
title: "PFTs"
output: html_document
date: "2026-03-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(lattice)
library(gstat)
library(stats)
library(sp)
library(ggplot2)
library(ggbreak)
library(scales)
library(tidyr)
library(plotly)
library(raster)
library(dismo)
library(sf)
library(tidyverse)
library(hrbrthemes)
library(terra)
library(RColorBrewer)
library(cowplot)
#PCA packages
library(corrplot)
library(factoextra) 
library(FactoMineR)
library(vegan)
library(mgcv)
library(ggrepel)
library(caret)
#netCDF packages
library(ncdf4)
library(CFtime)
library(stars)
library(here)
library(cowplot)
```


##Locations 
```{r}
data_loc <- here("Data", "TBS_Datasets")
pca_loc <- here("Data", "PCA")
outputs_loc <- here("Outputs")
averages_loc <- here("Outputs", "AvgRasters")
modelsNat_loc <- here("Outputs", "Models_Nat")
bnfrasters_loc <- here("Outputs", "AvgRasters", "BNF")
rast_loc <- here("Data", "ClippedGloballyGridded")
figure_loc <- here("Figures")

model_nat_figure <- here("Figures", "Model_Natural")
modeldata <- here("Data", "TRENDYN")

ebf_loc <- here("Data", "ClippedGloballyGridded", "EBF")
arid_loc <- here("Data", "ClippedGloballyGridded", "Arid")
natural_loc <- here("Data", "ClippedGloballyGridded", "Natural")

landcover_loc <- here("Data", "LandCover")
```




##CLASSIC

```{r}
Classic_landCoverFrac <- ncdf4::nc_open(here(landcover_loc, "CLASSIC_S0_landCoverFrac.nc"))
Classic_landCoverFrac
```
```{r}
names(Classic_landCoverFrac$var)
```

###Find the structire within our variable of interest. In this case it's longitude (360) × latitude (180) × PFT (10) × time (320)
```{r}
Classic_landCoverFrac$var[["landCoverFrac"]]$dim
```

###Extract only PFT 3 (EBF)
```{r}
var <- ncvar_get(Classic_landCoverFrac, "landCoverFrac",
                 start = c(1, 1, 3, 1),
                 count = c(-1, -1, 1, -1))
```

###Extract our dates of interest, 2000 to 2020
```{r}
time <- ncvar_get(Classic_landCoverFrac, "time")
dates <- as.Date(time, origin = "1700-12-31")
```

```{r}
time_idx <- which(year(dates) >= 2000 & year(dates) <= 2020)
var_subset <- var[ , , time_idx]
```

###Average fraction values over time
```{r}
mean_pft3 <- apply(var_subset, c(1, 2), mean, na.rm = TRUE)
```

###Rasterize
```{r}
lon <- ncvar_get(Classic_landCoverFrac, "longitude")
lat <- ncvar_get(Classic_landCoverFrac, "latitude")

r <- rast(t(mean_pft3))
ext(r) <- c(min(lon), max(lon), min(lat), max(lat))
crs(r) <- "EPSG:4326"

# Fix orientation (very likely needed)
r <- flip(r, direction = "vertical")

plot(r)
```


###Convert to areas:
```{r}
# area of each pixel in km²
#Terra accounts for the curvature of the earth doing it this way
cell_area <- cellSize(r, unit = "km")
```

```{r}
pft3_area <- r * cell_area
```

```{r}
plot(pft3_area)
```

###Check to make sure fraction is 0-1
```{r}
minmax(r)
```

```{r}
output_file <- here(outputs_loc,"PFTMaps", "Classic_EBFPFT.tif")

writeRaster(pft3_area, output_file, overwrite = TRUE)
```

###Total area
```{r}
total_area_km2 <- global(pft3_area, fun = "sum", na.rm = TRUE)
total_area_km2
```

###Total area of the tropics

```{r}
lat_r <- init(r, "y")
```

```{r}
tropics <- lat_r >= -23.5 & lat_r <= 23.5
```

```{r}
r2 <- r[[1]]   # or mean/any single timestep
land_mask <- r2 > 0
```

```{r}
cell_area <- cellSize(r2, unit = "km")
```

```{r}
tropical_land_area <- cell_area * tropics * land_mask
```

```{r}
total <- global(tropical_land_area, "sum", na.rm = TRUE)
total
```


###Find the area of the productive tropics - from the GPP mask 
```{r}
NPP_nat <- rast("/Users/katienelson/Documents/GitHub/TropicalBlindspots/Data/ClippedGloballyGridded/Natural/NPP_Nat_test.tif")
```

```{r}
cell_area <- cellSize(NPP_nat, unit = "km")

# Keep only cells with non-zero values
nonzero_area <- ifel(NPP_nat != 0 & !is.na(NPP_nat), cell_area, NA)

# Sum total area
total_area_km2 <- global(nonzero_area, "sum", na.rm = TRUE)

print(total_area_km2)
```












##CLM EBF

```{r}
CLM_landCoverFrac <- ncdf4::nc_open(here(landcover_loc, "CLM5.0_S0_landCoverFrac_PFT04.nc"))
CLM_landCoverFrac
```

```{r}
names(CLM_landCoverFrac$var)
```

###Find the structire within our variable of interest. In this case it's longitude (360) × latitude (180) × PFT (10) × time (320)
```{r}
CLM_landCoverFrac$var[["landCoverFrac"]]$dim
```

###Dont need to extract a PFT, already done within the file

```{r}
time <- ncvar_get(CLM_landCoverFrac, "time")
dates <- as.Date(time, origin = "1701-01-01")

time_idx <- which(lubridate::year(dates) >= 2000 &
                  lubridate::year(dates) <= 2020)

var_subset <- var[time_idx, , ]
```


###Average fraction values over time
```{r}
mean_map <- apply(var_subset, c(2,3), mean, na.rm = TRUE)
```

###Rasterize
```{r}
lon <- ncvar_get(CLM_landCoverFrac, "lon")
lat <- ncvar_get(CLM_landCoverFrac, "lat")

r <- rast(t(mean_map))
ext(r) <- c(min(lon), max(lon), min(lat), max(lat))
crs(r) <- "EPSG:4326"

# Fix orientation (very likely needed)
r <- flip(r, direction = "vertical")

plot(r)
```


###Convert to areas:
```{r}
# area of each pixel in km²
#Terra accounts for the curvature of the earth doing it this way
cell_area <- cellSize(r, unit = "km")
```

```{r}
pft_area <- r * cell_area
```

```{r}
plot(pft_area)
```

###Check to make sure fraction is 0-1
```{r}
minmax(r)
```

```{r}
output_file <- here(outputs_loc,"PFTMaps", "CLM5_EBFPFT.tif")

writeRaster(pft_area, output_file, overwrite = TRUE)
```

###Total area
```{r}
total_area_km2 <- global(pft_area, fun = "sum", na.rm = TRUE)
total_area_km2
```


###CLM EDF OVERWRITES

```{r}
CLM_landCoverFrac <- ncdf4::nc_open(here(landcover_loc, "CLM5.0_S0_landCoverFrac_PFT06.nc"))
CLM_landCoverFrac
```

```{r}
names(CLM_landCoverFrac$var)
```

###Find the structire within our variable of interest. In this case it's longitude (360) × latitude (180) × PFT (10) × time (320)
```{r}
CLM_landCoverFrac$var[["landCoverFrac"]]$dim
```

###Dont need to extract a PFT, already done within the file

```{r}
time <- ncvar_get(CLM_landCoverFrac, "time")
dates <- as.Date(time, origin = "1701-01-01")

time_idx <- which(lubridate::year(dates) >= 2000 &
                  lubridate::year(dates) <= 2020)

var_subset <- var[time_idx, , ]
```


###Average fraction values over time
```{r}
mean_map <- apply(var_subset, c(2,3), mean, na.rm = TRUE)
```

###Rasterize
```{r}
lon <- ncvar_get(CLM_landCoverFrac, "lon")
lat <- ncvar_get(CLM_landCoverFrac, "lat")

r <- rast(t(mean_map))
ext(r) <- c(min(lon), max(lon), min(lat), max(lat))
crs(r) <- "EPSG:4326"

# Fix orientation (very likely needed)
r <- flip(r, direction = "vertical")

plot(r)
```


###Convert to areas:
```{r}
# area of each pixel in km²
#Terra accounts for the curvature of the earth doing it this way
cell_area <- cellSize(r, unit = "km")
```

```{r}
pft_area <- r * cell_area
```

```{r}
plot(pft_area)
```

###Check to make sure fraction is 0-1
```{r}
minmax(r)
```

```{r}
output_file <- here(outputs_loc,"PFTMaps", "CLM5_EDFPFT.tif")

writeRaster(pft_area, output_file, overwrite = TRUE)
```



```{r}
total_area_km2 <- global(pft_area, fun = "sum", na.rm = TRUE)
total_area_km2
```









##Cable-Pop

```{r}
CP_landCoverFrac <- ncdf4::nc_open(here(landcover_loc, "CABLE-POP_S0_landCoverFrac.nc"))
CP_landCoverFrac
```

```{r}
names(CP_landCoverFrac$var)
```

###Find the structire within our variable of interest. In this case it's longitude (360) × latitude (180) × PFT (10) × time (320)
```{r}
CP_landCoverFrac$var[["landCoverFrac"]]$dim
```

###Extract only PFT 1 (EBF)
```{r}
#Seems like the actual value for EBT is in PFT 2?? 
var <- ncvar_get(CP_landCoverFrac, "landCoverFrac",
                 start = c(1, 1, 2),
                 count = c(-1, -1, 1))
```

###This PFT does not have time

###Rasterize
```{r}
lon <- ncvar_get(CP_landCoverFrac, "longitude")
lat <- ncvar_get(CP_landCoverFrac, "latitude")

r <- rast(t(var))
ext(r) <- c(min(lon), max(lon), min(lat), max(lat))
crs(r) <- "EPSG:4326"

# Fix orientation (very likely needed)
r <- flip(r, direction = "vertical")

plot(r)
```


###Convert to areas:
```{r}
# area of each pixel in km²
#Terra accounts for the curvature of the earth doing it this way
cell_area <- cellSize(r, unit = "km")
```

```{r}
var_area <- r * cell_area
```

```{r}
plot(var_area)
```

###Check to make sure fraction is 0-1
```{r}
minmax(r)
```

```{r}
output_file <- here(outputs_loc,"PFTMaps", "Cable-Pop_EBFPFT.tif")

writeRaster(var_area, output_file, overwrite = TRUE)
```

###Total area
```{r}
total_area_km2 <- global(var_area, fun = "sum", na.rm = TRUE)
total_area_km2
```









##JSBACH

```{r}
JSB_landCoverFrac <- ncdf4::nc_open(here(landcover_loc, "JSBACH_S0_landCoverFrac.nc"))
JSB_landCoverFrac
```

```{r}
names(JSB_landCoverFrac$var)
```

###Find the structire within our variable of interest. In this case it's longitude (360) × latitude (180) × PFT (10) × time (320)
```{r}
JSB_landCoverFrac$var[["landCoverFrac"]]$dim
```

###Extract only PFT 1 (EBF)
```{r}
#Seems like the actual value for EBT is in PFT 2?? 
var <- ncvar_get(JSB_landCoverFrac, "landCoverFrac",
                 start = c(1, 1, 3),
                 count = c(-1, -1, 1))
```

###This PFT does not have time

###Rasterize
```{r}
lon <- ncvar_get(JSB_landCoverFrac, "longitude")
lat <- ncvar_get(JSB_landCoverFrac, "latitude")

r <- rast(t(var))
ext(r) <- c(min(lon), max(lon), min(lat), max(lat))
crs(r) <- "EPSG:4326"

# Fix orientation (very likely needed)
r <- flip(r, direction = "vertical")

plot(r)
```


###Convert to areas:
```{r}
# area of each pixel in km²
#Terra accounts for the curvature of the earth doing it this way
cell_area <- cellSize(r, unit = "km")
```

```{r}
var_area <- r * cell_area
```

```{r}
plot(var_area)
```

###Check to make sure fraction is 0-1
```{r}
minmax(r)
```

```{r}
output_file <- here(outputs_loc,"PFTMaps", "JSBach_EBFPFT.tif")

writeRaster(var_area, output_file, overwrite = TRUE)
```


###Total area
```{r}
total_area_km2 <- global(var_area, fun = "sum", na.rm = TRUE)
total_area_km2
```







##JULES

```{r}
JULES_landCoverFrac <- ncdf4::nc_open(here(landcover_loc, "JULES_S0_landCoverFrac.nc"))
JULES_landCoverFrac
```

```{r}
names(JULES_landCoverFrac$var)
```

###Find the structire within our variable of interest. In this case it's longitude (360) × latitude (180) × PFT (10) × time (320)
```{r}
JULES_landCoverFrac$var[["landCoverFrac"]]$dim
```

###Extract only PFT 1 (EBF)
```{r}
#Seems like the actual value for EBT is in PFT 2?? 
var <- ncvar_get(JULES_landCoverFrac, "landCoverFrac",
                 start = c(1, 1, 2, 1),
                 count = c(-1, -1, 1, -1))
```

###Extract our dates of interest, 2000 to 2020
```{r}
time <- ncvar_get(JULES_landCoverFrac, "time")
dates <- as.Date(time, origin = "1700-12-31")
```

```{r}
time_idx <- which(year(dates) >= 2000 & year(dates) <= 2020)
var_subset <- var[ , , time_idx]
```

###Average fraction values over time
```{r}
mean_pft <- apply(var_subset, c(1, 2), mean, na.rm = TRUE)
```

###Rasterize
```{r}
lon <- ncvar_get(JULES_landCoverFrac, "lon")
lat <- ncvar_get(JULES_landCoverFrac, "lat")

lon[lon > 180] <- lon[lon > 180] - 360

# get ordering index
ord <- order(lon)
# reorder lon
lon <- lon[ord]
# reorder your data (IMPORTANT)
mean_pft <- mean_pft[ord, ]

r <- rast(t(mean_pft))
ext(r) <- c(min(lon), max(lon), min(lat), max(lat))
crs(r) <- "EPSG:4326"

# Fix orientation (very likely needed)
r <- flip(r, direction = "vertical")

plot(r)
```


###Convert to areas:
```{r}
# area of each pixel in km²
#Terra accounts for the curvature of the earth doing it this way
cell_area <- cellSize(r, unit = "km")
```

```{r}
var_area <- r * cell_area
```

```{r}
plot(var_area)
```

###Check to make sure fraction is 0-1
```{r}
minmax(r)
```

```{r}
output_file <- here(outputs_loc,"PFTMaps", "JULES_EBFPFT.tif")

writeRaster(var_area, output_file, overwrite = TRUE)
```


###Total area
```{r}
total_area_km2 <- global(var_area, fun = "sum", na.rm = TRUE)
total_area_km2
```












##LPJ-Guess

```{r}
LPJ_landCoverFrac <- ncdf4::nc_open(here(landcover_loc, "LPJ-GUESS_S0_landCoverFrac.nc"))
LPJ_landCoverFrac
```

```{r}
names(LPJ_landCoverFrac$var)
```

###Find the structire within our variable of interest. In this case it's longitude (360) × latitude (180) × PFT (10) × time (320)
```{r}
LPJ_landCoverFrac$var[["landCoverFrac"]]$dim
```

###Extract PFTs 8,9,10
```{r}
pft8 <- ncvar_get(LPJ_landCoverFrac, "landCoverFrac",
                  start = c(8, 1, 1, 1),
                  count = c(1, -1, -1, -1))

pft9 <- ncvar_get(LPJ_landCoverFrac, "landCoverFrac",
                  start = c(9, 1, 1, 1),
                  count = c(1, -1, -1, -1))

pft10 <- ncvar_get(LPJ_landCoverFrac, "landCoverFrac",
                   start = c(10, 1, 1, 1),
                   count = c(1, -1, -1, -1))

pft_combined <- pft8 + pft9 + pft10
```

###Extract our dates of interest, 2000 to 2020
```{r}
time <- ncvar_get(LPJ_landCoverFrac, "time")
dates <- as.Date(time, origin = "1700-12-31")
```

```{r}
time_idx <- which(year(dates) >= 2000 & year(dates) <= 2020)
pft_subset <- pft_combined[ , , time_idx]
```

###Average fraction values over time
```{r}
mean_pft <- apply(pft_subset, c(1, 2), mean, na.rm = TRUE)
```

###Rasterize
```{r}
lon <- ncvar_get(LPJ_landCoverFrac, "longitude")
lat <- ncvar_get(LPJ_landCoverFrac, "latitude")

lon[lon > 180] <- lon[lon > 180] - 360

# get ordering index
ord <- order(lon)
# reorder lon
lon <- lon[ord]
# reorder your data (IMPORTANT)
mean_pft <- mean_pft[ord, ]

r <- rast(t(mean_pft))
ext(r) <- c(min(lon), max(lon), min(lat), max(lat))
crs(r) <- "EPSG:4326"

# Fix orientation (very likely needed)
r <- flip(r, direction = "vertical")

plot(r)
```


###Convert to areas:
```{r}
# area of each pixel in km²
#Terra accounts for the curvature of the earth doing it this way
cell_area <- cellSize(r, unit = "km")
```

```{r}
var_area <- r * cell_area
```

```{r}
plot(var_area)
```

###Check to make sure fraction is 0-1
```{r}
minmax(r)
```

```{r}
output_file <- here(outputs_loc,"PFTMaps", "LPJ-Guess_EBFPFT.tif")

writeRaster(var_area, output_file, overwrite = TRUE)
```


###Total area
```{r}
total_area_km2 <- global(var_area, fun = "sum", na.rm = TRUE)
total_area_km2
```





##LPX-Bern

```{r}
LPXB_landCoverFrac <- ncdf4::nc_open(here(landcover_loc, "LPX-Bern_S0_landCoverFrac.nc"))
LPXB_landCoverFrac
```



```{r}
names(LPXB_landCoverFrac$var)
```

###Find the structire within our variable of interest. In this case it's longitude (360) × latitude (180) × PFT (10) × time (320)
```{r}
LPXB_landCoverFrac$var[["landCoverFrac"]]$dim
```

###Extract PFTs and average time steps in one (files too big)
```{r}
time <- ncvar_get(LPXB_landCoverFrac, "time")
dates <- as.Date(time, origin = "1700-12-31")

time_idx <- which(lubridate::year(dates) >= 2000 & lubridate::year(dates) <= 2020)

pft_mean <- 0

for (p in c(1, 2, 13, 16, 17)) {
  tmp <- ncvar_get(LPXB_landCoverFrac, "landCoverFrac",
                   start = c(1, 1, p, min(time_idx)),
                   count = c(-1, -1, 1, length(time_idx)))

  tmp_mean <- apply(tmp, c(1, 2), mean, na.rm = TRUE)
  
  pft_mean <- pft_mean + tmp_mean
  
  rm(tmp, tmp_mean)
  gc()
}
```

###Rasterize
```{r}
lon <- ncvar_get(LPXB_landCoverFrac, "longitude")
lat <- ncvar_get(LPXB_landCoverFrac, "latitude")

lon[lon > 180] <- lon[lon > 180] - 360

# get ordering index
ord <- order(lon)
# reorder lon
lon <- lon[ord]
# reorder your data (IMPORTANT)
pft_mean <- pft_mean[ord, ]

r <- rast(t(pft_mean))
ext(r) <- c(min(lon), max(lon), min(lat), max(lat))
crs(r) <- "EPSG:4326"

# Fix orientation (very likely needed)
r <- flip(r, direction = "vertical")

plot(r)
```


###Convert to areas:
```{r}
# area of each pixel in km²
#Terra accounts for the curvature of the earth doing it this way
cell_area <- cellSize(r, unit = "km")
```

```{r}
var_area <- r * cell_area
```

```{r}
plot(var_area)
```

###Check to make sure fraction is 0-1
```{r}
minmax(r)
```

```{r}
output_file <- here(outputs_loc,"PFTMaps", "LPX-Bern_EBFPFT.tif")

writeRaster(var_area, output_file, overwrite = TRUE)
```


###Total area
```{r}
total_area_km2 <- global(var_area, fun = "sum", na.rm = TRUE)
total_area_km2
```






##Orchidee

```{r}
ORC_landCoverFrac <- ncdf4::nc_open(here(landcover_loc, "ORCHIDEE_S0_landCoverFrac.nc"))
ORC_landCoverFrac
```



```{r}
names(ORC_landCoverFrac$var)
```

###Find the structire within our variable of interest. In this case it's longitude (360) × latitude (180) × PFT (10) × time (320)
```{r}
ORC_landCoverFrac$var[["landCoverFrac"]]$dim
```

###Extract PFTs and average time steps in one (files too big)
```{r}
time <- ncvar_get(ORC_landCoverFrac, "time")
dates <- as.Date(time, origin = "1700-12-31")

time_idx <- which(lubridate::year(dates) >= 2000 & lubridate::year(dates) <= 2020)

pft_mean <- 0

for (p in c(2,3)) {
  tmp <- ncvar_get(ORC_landCoverFrac, "landCoverFrac",
                   start = c(1, 1, p, min(time_idx)),
                   count = c(-1, -1, 1, length(time_idx)))

  tmp_mean <- apply(tmp, c(1, 2), mean, na.rm = TRUE)
  
  pft_mean <- pft_mean + tmp_mean
  
  rm(tmp, tmp_mean)
  gc()
}
```

###Rasterize
```{r}
lon <- ncvar_get(ORC_landCoverFrac, "longitude")
lat <- ncvar_get(ORC_landCoverFrac, "latitude")

lon[lon > 180] <- lon[lon > 180] - 360

# get ordering index
ord <- order(lon)
# reorder lon
lon <- lon[ord]
# reorder your data (IMPORTANT)
pft_mean <- pft_mean[ord, ]

r <- rast(t(pft_mean))
ext(r) <- c(min(lon), max(lon), min(lat), max(lat))
crs(r) <- "EPSG:4326"

# Fix orientation (very likely needed)
r <- flip(r, direction = "vertical")

plot(r)
```


###Convert to areas:
```{r}
# area of each pixel in km²
#Terra accounts for the curvature of the earth doing it this way
cell_area <- cellSize(r, unit = "km")
```

```{r}
var_area <- r * cell_area
```

```{r}
plot(var_area)
```

###Check to make sure fraction is 0-1
```{r}
minmax(r)
```

```{r}
output_file <- here(outputs_loc,"PFTMaps", "Orchidee_EBFPFT.tif")

writeRaster(var_area, output_file, overwrite = TRUE)
```



###Total area
```{r}
total_area_km2 <- global(var_area, fun = "sum", na.rm = TRUE)
total_area_km2
```













##OCN

```{r}
OCN_landCoverFrac <- ncdf4::nc_open(here(landcover_loc, "OCN_S0_landCoverFrac.nc"))
OCN_landCoverFrac
```

```{r}
names(OCN_landCoverFrac$var)
```

###Find the structire within our variable of interest. In this case it's longitude (360) × latitude (180) × PFT (10) × time (320)
```{r}
OCN_landCoverFrac$var[["landCoverFrac"]]$dim
```

###Extract PFTs and average time steps in one (files too big) long, lat, pft, time
```{r}
time <- ncvar_get(OCN_landCoverFrac, "time")
dates <- as.Date(time, origin = "1700-12-31")

time_idx <- which(lubridate::year(dates) >= 2000 & lubridate::year(dates) <= 2020)

pft_mean <- 0

for (p in c(2,3)) {
  tmp <- ncvar_get(OCN_landCoverFrac, "landCoverFrac",
                   start = c(1, 1, p, min(time_idx)),
                   count = c(-1, -1, 1, length(time_idx)))

  tmp_mean <- apply(tmp, c(1, 2), mean, na.rm = TRUE)
  
  pft_mean <- pft_mean + tmp_mean
  
  rm(tmp, tmp_mean)
  gc()
}
```

###Rasterize
```{r}
lon <- ncvar_get(OCN_landCoverFrac, "longitude")
lat <- ncvar_get(OCN_landCoverFrac, "latitude")

lon[lon > 180] <- lon[lon > 180] - 360

# get ordering index
ord <- order(lon)
# reorder lon
lon <- lon[ord]
# reorder your data (IMPORTANT)
pft_mean <- pft_mean[ord, ]

r <- rast(t(pft_mean))
ext(r) <- c(min(lon), max(lon), min(lat), max(lat))
crs(r) <- "EPSG:4326"

# Fix orientation (very likely needed)
r <- flip(r, direction = "vertical")

plot(r)
```


###Convert to areas:
```{r}
# area of each pixel in km²
#Terra accounts for the curvature of the earth doing it this way
cell_area <- cellSize(r, unit = "km")
```

```{r}
var_area <- r * cell_area
```

```{r}
plot(var_area)
```

###Check to make sure fraction is 0-1
```{r}
minmax(r)
```

```{r}
output_file <- here(outputs_loc,"PFTMaps", "OCN_EBFPFT.tif")

writeRaster(var_area, output_file, overwrite = TRUE)
```



###Total area
```{r}
total_area_km2 <- global(var_area, fun = "sum", na.rm = TRUE)
total_area_km2
```









##ISAM
```{r}
ISAM_landCoverFrac <- ncdf4::nc_open(here(landcover_loc, "ISAM_S0_landCoverFrac.nc"))
ISAM_landCoverFrac
```



```{r}
names(ISAM_landCoverFrac$var)
```

###Find the structire within our variable of interest. In this case it's longitude (360) × latitude (180) × PFT (10) × time (320)
```{r}
ISAM_landCoverFrac$var[["landCoverFrac"]]$dim
```

###Extract PFTs and average time steps in one (files too big) long, lat, pft, time
###This model has an extra class for year and is only a value of one, plus the dates are by year
```{r}
time <- ncvar_get(ISAM_landCoverFrac, "time")
time_idx <- which(time >= 2000 & time <= 2020)


pft_mean <- 0

for (p in c(1,2)) {
  tmp <- ncvar_get(ISAM_landCoverFrac, "landCoverFrac",
                   start = c(1, 1, p, 1, min(time_idx)),
                   count = c(-1, -1, 1, 1, length(time_idx)))

  tmp_mean <- apply(tmp, c(1, 2), mean, na.rm = TRUE)
  
  pft_mean <- pft_mean + tmp_mean
  
  rm(tmp, tmp_mean)
  gc()
}
```




###Rasterize
```{r}
lon <- ncvar_get(ISAM_landCoverFrac, "longitude")
lat <- ncvar_get(ISAM_landCoverFrac, "latitude")

lon[lon > 180] <- lon[lon > 180] - 360

# get ordering index
ord <- order(lon)
# reorder lon
lon <- lon[ord]
# reorder your data (IMPORTANT)
pft_mean <- pft_mean[ord, ]

r <- rast(t(pft_mean))
ext(r) <- c(min(lon), max(lon), min(lat), max(lat))
crs(r) <- "EPSG:4326"

# Fix orientation (very likely needed)
r <- flip(r, direction = "vertical")

plot(r)
```


###Convert to areas:
```{r}
# area of each pixel in km²
#Terra accounts for the curvature of the earth doing it this way
cell_area <- cellSize(r, unit = "km")
```

```{r}
var_area <- r * cell_area
```

```{r}
plot(var_area)
```

###Check to make sure fraction is 0-1
```{r}
minmax(r)
```

```{r}
output_file <- here(outputs_loc,"PFTMaps", "ISAM_EBFPFT.tif")

writeRaster(var_area, output_file, overwrite = TRUE)
```



###Total area
```{r}
total_area_km2 <- global(var_area, fun = "sum", na.rm = TRUE)
total_area_km2
```









