---
title: "TropicalBlindSpotsCode"
output: html_document
date: "2025-05-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(lattice)
library(gstat)
library(stats)
library(sp)
library(ggplot2)
library(ggbreak)
library(scales)
library(tidyr)
library(plotly)
library(raster)
library(dismo)
library(sf)
library(tidyverse)
library(hrbrthemes)
library(terra)
library(RColorBrewer)
library(cowplot)
#PCA packages
library(corrplot)
library(factoextra) 
library(FactoMineR)
library(vegan)
library(mgcv)
library(ggrepel)
library(caret)
#netCDF packages
library(ncdf4)
library(CFtime)
library(stars)
library(here)
library(cowplot)
```

##Locations 
```{r}
data_loc <- here("Data", "TBS_Datasets")
pca_loc <- here("Data", "PCA")
outputs_loc <- here("Outputs")
averages_loc <- here("Outputs", "AvgRasters")
modelsNat_loc <- here("Outputs", "Models_Nat")
bnfrasters_loc <- here("Outputs", "AvgRasters", "BNF")
rast_loc <- here("Data", "ClippedGloballyGridded")
figure_loc <- here("Figures")
model_nat_figure <- here("Figures", "Model_Natural")
modeldata <- here("Data", "TRENDYN")
ebf_loc <- here("Data", "ClippedGloballyGridded", "EBF")
arid_loc <- here("Data", "ClippedGloballyGridded", "Arid")
natural_loc <- here("Data", "ClippedGloballyGridded", "Natural")
```

##Data files 
```{r}
bnf <- read.csv(here(data_loc,'BNF_Sampled.csv'), header= TRUE)
denit <- read.csv(here(data_loc,'Denitrification_Sampled.csv'), header= TRUE)
```

```{r}
bnf.ebf <- bnf %>% filter(Biome_rough == "EBF")
denit.ebf <- denit %>% filter(Biome_rough == "EBF")

bnf.arid <- bnf %>%
  filter(Biome_rough %in% c("Savannas", "Woody savannas", "Grassland"))
denit.arid <- denit %>%
  filter(Biome_rough %in% c("Savannas", "Woody savannas", "Grassland"))
```

###Expand to include repeat sampling
```{r}
bnf_expanded <- bnf %>%
  uncount(weights = Number_of_sites)
```

```{r}
bnf.ebf.ex <- bnf_expanded %>% filter(Biome_rough == "EBF")

bnf.arid.ex <- bnf_expanded %>%
  filter(Biome_rough %in% c("Savannas", "Woody savannas", "Grassland"))
```




##IGBP data

```{r}
igbp_rough_sample <- read.csv(here(data_loc, 'IGBP_Plot.csv'), header = TRUE)
igbp_rough_repeats <- read.csv(here(data_loc, 'IGBP_Plot_repeats.csv'), header = TRUE)
```





###IGBP calculations with repeats 
```{r}
bnf %>%
  group_by(Biome_rough) %>%
  summarise(total_samples = sum(Number_of_sites, na.rm = TRUE))
```


```{r}
denit %>%
  group_by(Biome_rough) %>%
  summarise(total_samples = sum(number_sites, na.rm = TRUE))
```


###IGBP Bubble chart
```{r}
df_long <- igbp_rough_repeats %>%
  pivot_longer(
    cols = c(Area_percent, Denit_percent, BNF_percent),
    names_to = "Process",
    values_to = "Percent"
  ) %>%
  # Optional: make nicer names for the legend
  mutate(Process = recode(Process,
                          "Area_percent" = "Area",
                          "Denit_percent" = "Denitrification",
                          "BNF_percent" = "BNF"))
```



```{r}
max_percent <- df_long %>%
  filter(Process %in% c("Denitrification", "BNF", "Area")) %>%
  summarise(max_val = max(Percent, na.rm = TRUE)) %>%
  pull(max_val)
```


```{r}
##Denitrification only 
df_plot <- df_long %>%
  filter(Process %in% c("Area", "Denitrification")) %>%  # ignore BNF
  mutate(
    y_pos = 1  # fixed y position for all bubbles
  )
custom_order_pretty <- c(
  "Savannas\nand \nShrublands",
  "Grasslands",
  "Evergreen \nBroadleaf \nForests",
  "Deciduous \nBroadleaf \nForests",
  "Barren",
  "Mixed\nForests",
  "Permanent \nWetlands\nand Coastal"

)
df_plot <- df_plot %>%
  mutate(Biome = factor(Biome, levels = c(
    "Savannas and Shrublands", "Grassland", "EBF", "DBF", "Barren", "Mixed Forests", "Permanent Wetlands and Coastal
"
  )))

legend_breaks <- c(1, 15, 30)

bubbles <- ggplot(df_plot, aes(x = Biome, size = Percent)) +

  # 1️⃣ Denitrification FIRST (background)
  geom_point(
    data = df_plot %>% filter(Process == "Denitrification"),
    aes(y = 0.85),
    color = "blue",
    alpha = 0.7
  ) +

  # 2️⃣ Area SECOND (on top, outline only)
  geom_point(
    data = df_plot %>% filter(Process == "Area"),
    aes(y = 0.85),
    shape = 21,          # circle that supports fill + outline
    fill = NA,           # no fill
    color = "grey",    # grey outline
    stroke = 1.2,        # outline thickness
    alpha = 1
  ) +
  scale_size_area(
    max_size = 25,
    limits = c(0, max_percent),  # <-- same as Denitrification
    breaks = c(1, 15, 30),
    labels = function(x) paste0(x, "%")
  ) +
  scale_color_manual(values = c(
    "Area" = "grey",
    "Denitrification" = "blue"
  )) +
  scale_x_discrete(labels = custom_order_pretty) +
  scale_y_continuous(
    breaks = NULL,  # hide y-axis
    limits = c(0, max(df_plot$Percent) * 0.1)  # set upper limit slightly above bubbles
  ) +
  theme_minimal() +
  theme(
    axis.title.y = element_blank(),
    panel.grid = element_blank(), 
  legend.position.inside = c(0.5, -0.05),  # x, y
    legend.justification = c(0.5, 0)
  ) +
  labs(size = "Proportional area \nor data coverage", color = "Process") +
  guides(
  size = guide_legend(
    override.aes = list(
      shape = 21,
      fill = NA,
      color = "grey",
      stroke = 1.2
    )
  )
)

bubbles
#ggsave(here(figure_loc, "IGBP_Bubbles_denit.png"), plot = bubbles, width = 8, height = 6, dpi = 300)
```



```{r}
#BNF only 
df_plot_bnf <- df_long %>%
  filter(Process %in% c("Area", "BNF")) %>%  
  mutate(
    y_pos = 1  # fixed y position for all bubbles
  )

#custom_order_pretty <- c(
 # "Savannas\nand \nShrublands \n(37.1%)",
#  "Grassland \n(29.4%)",
 # "Evergreen \nBroadleaf \nForests \n(28.1%)",
  #"Deciduous \nBroadleaf \nForests \n(2.2%)",
  #"Barren \n(1.3%)",
  #"Mixed\nForests \n(1.1%)",
  #"Permanent \nWetlands\nand Coastal \n(0.7%)"

#)
df_plot_bnf <- df_plot_bnf %>%
  mutate(Biome = factor(Biome, levels = c(
    "Savannas and Shrublands", "Grassland", "EBF", "DBF", "Barren", "Mixed Forests", "Permanent Wetlands and Coastal
"
  )))

legend_breaks <- c(1, 25, 50)

bubbles <- ggplot(df_plot_bnf, aes(x = Biome, size = Percent, color = Process)) +
  # Plot "Area" first (underneath)
  # 1️⃣ Denitrification FIRST (background)
  geom_point(
    data = df_plot_bnf %>% filter(Process == "BNF"),
    aes(y = 0.85),
    color = "red",
    alpha = 0.7
  ) +

  # 2️⃣ Area SECOND (on top, outline only)
  geom_point(
    data = df_plot_bnf %>% filter(Process == "Area"),
    aes(y = 0.85),
    shape = 21,          # circle that supports fill + outline
    fill = NA,           # no fill
    color = "grey",    # grey outline
    stroke = 1.2,        # outline thickness
    alpha = 1
  ) +
  
  scale_size_area(
    max_size = 25,
    limits = c(0, max_percent),  # <-- same as Denitrification
    breaks = c(1, 15, 30),
    labels = function(x) paste0(x, "%")
  ) +
  scale_color_manual(values = c(
    "Area" = "grey",
    "BNF" = "red"
  )) +
  scale_x_discrete(labels = custom_order_pretty) +
  scale_y_continuous(
    breaks = NULL,  # hide y-axis
    limits = c(0, max(df_plot$Percent) * 0.1)  # set upper limit slightly above bubbles
  ) +
  theme_minimal() +
  theme(
    axis.title.y = element_blank(),
    panel.grid = element_blank()  # remove all grid lines
  ) +
  labs(size = "Percent", color = "Process") +
  guides(
  size = guide_legend(
    override.aes = list(
      shape = 21,
      fill = NA,
      color = "grey",
      stroke = 1.2
    )
  )
)

bubbles
#ggsave(here(figure_loc, "IGBP_Bubbles_BNF_expanded.png"), plot = bubbles, width = 8, height = 6, dpi = 300)
```






###IGBP Bubbles repeats
```{r}
df_long <- igbp_rough_repeats %>%
  pivot_longer(
    cols = c(Area_percent, Denit_percent, BNF_percent),
    names_to = "Process",
    values_to = "Percent"
  ) %>%
  # Optional: make nicer names for the legend
  mutate(Process = recode(Process,
                          "Area_percent" = "Area",
                          "Denit_percent" = "Denitrification",
                          "BNF_percent" = "BNF"))
```


```{r}
custom_order_pretty <- c(
  "Savannas\nand \nShrublands \n(37.1%)",
  "Grassland \n(29.4%)",
  "Evergreen \nBroadleaf \nForests \n(28.1%)",
  "Deciduous \nBroadleaf \nForests \n(2.2%)",
  "Barren \n(1.3%)",
  "Mixed\nForests \n(1.1%)",
  "Permanent \nWetlands\nand Coastal \n(0.7%)"

)
df_long <- df_long %>%
  mutate(Biome = factor(Biome, levels = c(
    "Savannas and Shrublands", "Grassland", "EBF", "DBF", "Barren", "Mixed Forests", "Permanent Wetlands and Coastal
"
  )))
# Define the representative values for the legend
legend_breaks <- c(min(1), 
                   median(30), 
                   max(50))
# Bubble plot
plot <- ggplot(df_long, aes(x = Biome, y = Process, size = Percent, color = Process)) +
  geom_point(alpha = 0.7) +
  scale_size_area(max_size = 25, breaks = legend_breaks) +
  scale_color_manual(values = c(
    "Area" = "grey",            
    "Denitrification" = "blue", 
    "BNF" = "red"               
  )) +
  scale_x_discrete(labels = custom_order_pretty) +
  theme_minimal() +
  labs(size = "Percent", color = "Process") +
  theme(axis.title.y = element_blank())

plot
#ggsave(here(figure_loc, "IGBP_Bubbles.png"), plot = plot, width = 8, height = 6, dpi = 300)
```

```{r}
max_percent <- df_long %>%
  filter(Process %in% c("Denitrification", "BNF", "Area")) %>%
  summarise(max_val = max(Percent, na.rm = TRUE)) %>%
  pull(max_val)
```


```{r}
##Denitrification only 
df_plot <- df_long %>%
  filter(Process %in% c("Area", "Denitrification")) %>%  # ignore BNF
  mutate(
    y_pos = 1  # fixed y position for all bubbles
  )
custom_order_pretty <- c(
  "Savannas\nand \nShrublands",
  "Grasslands",
  "Evergreen \nBroadleaf \nForests",
  "Deciduous \nBroadleaf \nForests",
  "Barren",
  "Mixed\nForests",
  "Permanent \nWetlands\nand Coastal"

)
df_plot <- df_plot %>%
  mutate(Biome = factor(Biome, levels = c(
    "Savannas and Shrublands", "Grassland", "EBF", "DBF", "Barren", "Mixed Forests", "Permanent Wetlands and Coastal
"
  )))

legend_breaks <- c(1, 15, 30)

bubbles <- ggplot(df_plot, aes(x = Biome, size = Percent)) +

  # 1️⃣ Denitrification FIRST (background)
  geom_point(
    data = df_plot %>% filter(Process == "Denitrification"),
    aes(y = 0.85),
    color = "blue",
    alpha = 0.7
  ) +

  # 2️⃣ Area SECOND (on top, outline only)
  geom_point(
    data = df_plot %>% filter(Process == "Area"),
    aes(y = 0.85),
    shape = 21,          # circle that supports fill + outline
    fill = NA,           # no fill
    color = "grey",    # grey outline
    stroke = 1.2,        # outline thickness
    alpha = 1
  ) +
  scale_size_area(
    max_size = 25,
    limits = c(0, max_percent),  # <-- same as Denitrification
    breaks = c(1, 15, 30),
    labels = function(x) paste0(x, "%")
  ) +
  scale_color_manual(values = c(
    "Area" = "grey",
    "Denitrification" = "blue"
  )) +
  scale_x_discrete(labels = custom_order_pretty) +
  scale_y_continuous(
    breaks = NULL,  # hide y-axis
    limits = c(0, max(df_plot$Percent) * 0.1)  # set upper limit slightly above bubbles
  ) +
  theme_minimal() +
  theme(
    axis.title.y = element_blank(),
    panel.grid = element_blank(), 
  legend.position.inside = c(0.5, -0.05),  # x, y
    legend.justification = c(0.5, 0)
  ) +
  labs(size = "Proportional area \nor data coverage", color = "Process") +
  guides(
  size = guide_legend(
    override.aes = list(
      shape = 21,
      fill = NA,
      color = "grey",
      stroke = 1.2
    )
  )
)

bubbles
#ggsave(here(figure_loc, "IGBP_Bubbles_denit.png"), plot = bubbles, width = 8, height = 6, dpi = 300)
```



```{r}
#BNF only 
df_plot_bnf <- df_long %>%
  filter(Process %in% c("Area", "BNF")) %>%  
  mutate(
    y_pos = 1  # fixed y position for all bubbles
  )

#custom_order_pretty <- c(
 # "Savannas\nand \nShrublands \n(37.1%)",
#  "Grassland \n(29.4%)",
 # "Evergreen \nBroadleaf \nForests \n(28.1%)",
  #"Deciduous \nBroadleaf \nForests \n(2.2%)",
  #"Barren \n(1.3%)",
  #"Mixed\nForests \n(1.1%)",
  #"Permanent \nWetlands\nand Coastal \n(0.7%)"

#)
df_plot_bnf <- df_plot_bnf %>%
  mutate(Biome = factor(Biome, levels = c(
    "Savannas and Shrublands", "Grassland", "EBF", "DBF", "Barren", "Mixed Forests", "Permanent Wetlands and Coastal
"
  )))

legend_breaks <- c(1, 25, 50)

bubbles <- ggplot(df_plot_bnf, aes(x = Biome, size = Percent, color = Process)) +
  # Plot "Area" first (underneath)
  # 1️⃣ Denitrification FIRST (background)
  geom_point(
    data = df_plot_bnf %>% filter(Process == "BNF"),
    aes(y = 0.85),
    color = "red",
    alpha = 0.7
  ) +

  # 2️⃣ Area SECOND (on top, outline only)
  geom_point(
    data = df_plot_bnf %>% filter(Process == "Area"),
    aes(y = 0.85),
    shape = 21,          # circle that supports fill + outline
    fill = NA,           # no fill
    color = "grey",    # grey outline
    stroke = 1.2,        # outline thickness
    alpha = 1
  ) +
  
  scale_size_area(
    max_size = 25,
    limits = c(0, max_percent),  # <-- same as Denitrification
    breaks = c(1, 15, 30),
    labels = function(x) paste0(x, "%")
  ) +
  scale_color_manual(values = c(
    "Area" = "grey",
    "BNF" = "red"
  )) +
  scale_x_discrete(labels = custom_order_pretty) +
  scale_y_continuous(
    breaks = NULL,  # hide y-axis
    limits = c(0, max(df_plot$Percent) * 0.1)  # set upper limit slightly above bubbles
  ) +
  theme_minimal() +
  theme(
    axis.title.y = element_blank(),
    panel.grid = element_blank()  # remove all grid lines
  ) +
  labs(size = "Percent", color = "Process") +
  guides(
  size = guide_legend(
    override.aes = list(
      shape = 21,
      fill = NA,
      color = "grey",
      stroke = 1.2
    )
  )
)

bubbles
#ggsave(here(figure_loc, "IGBP_Bubbles_BNF.png"), plot = bubbles, width = 8, height = 6, dpi = 300)
```














##Citations
```{r}
citations <- read.csv(here(data_loc, 'Citations_Region.csv'), header = TRUE)
```

```{r}
citations %>%
  group_by(Region) %>%
  summarise(n = n())
```

```{r}
df <- citations %>%
  group_by(Region) %>%
  summarise(
    count = n(),
    total_value = sum(citations, na.rm = TRUE)
  ) %>%
  mutate(citationspersite = total_value / count)
```

```{r}
ggplot(df, aes(x = Region, y = citationspersite)) +
  geom_bar(stat = "identity", fill = "black") +
  labs(x = "Region", y = "Citations normalized to number of studies") +
  theme_minimal()
```







##Density Plots
###Though shown here for phosphorous, code can be adapted for any single variable. 

Soil Phosphorous
```{r}
SoilP <- raster(here(rast_loc,'SoilP_NoSahara.tif'))

# plot raster data
plot(SoilP,
     main = "Soil P")
```


Convert raster values to data frame
```{r}
SoilP <- brick(here(rast_loc,'SoilP_NoSahara.tif'))
```

```{r}
SoilP_df <- as.data.frame(rasterToPoints(SoilP,spatial = TRUE))
head(SoilP_df)
hist(SoilP_df$SoilP_NoSahara)
```


Density curves with lines indicating the peak (mode) of each population
```{r}
median_tropics <- median(SoilP_df$SoilP_NoSahara, na.rm = TRUE)
median_bnf     <- median(bnf_expanded$soil_total_phosphorus_EX, na.rm = TRUE)
median_denit   <- median(denit$Extracted_soil_total_phosphorus, na.rm = TRUE)

# --- 2. Set axis limits and color palette ---
x_axis_extent <- c(0, 1250)

color_palette <- c(
  "Tropics" = "grey",
  "BNF data" = "red",
  "Denitrification \ndata" = "blue"
)

# --- 3. Create the plot with density curves and vertical dashed lines ---
p <- ggplot() +
  geom_density(data = SoilP_df,
      aes(x = SoilP_NoSahara,
        fill = factor("Tropics", levels = names(color_palette))),
    alpha = 0.35
  ) +
  geom_density(data = bnf_expanded, aes(x = soil_total_phosphorus_EX, fill = "BNF data"), alpha = 0.35) +
  geom_density(data = denit, aes(x = Extracted_soil_total_phosphorus, fill = "Denitrification \ndata"), alpha = 0.35) +
  
  # Add vertical dashed lines at density peaks (modes)
  geom_vline(xintercept = median_tropics, linetype = "dashed", color = "grey20", linewidth = 0.75) +
  geom_vline(xintercept = median_bnf,     linetype = "dashed", color = color_palette["BNF data"], linewidth = 0.75) +
  geom_vline(xintercept = median_denit,   linetype = "dashed", color = color_palette["Denitrification \ndata"], linewidth = 0.75) +

  # Labels, theme, and scale adjustments
  labs(title = "Tropics wide", x = expression("Soil P (mg kg"^{-1}*")"), y = "Density") +
  theme_minimal() +
  xlim(x_axis_extent) +
  scale_fill_manual(values = color_palette, name = " ") +
  theme(
    legend.position = "none",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.title = element_text(size = 20),
    axis.text = element_text(size = 18),
    axis.title.y = element_text(margin = margin(r = 15)),
    plot.title = element_text(size = 22, face = "bold", hjust = 0.5)  # Change font size and make it bold
  )

# Show plot
p
#ggsave(here(figure_loc, "MedianDensity/SoilP_NoSahara_Density_expanded.png"), plot = p, width = 8, height = 6, dpi = 400)
```



###Density Plots - EBF

```{r}
SoilP <- raster(here(ebf_loc,'SoilP_EBFOnly.tif'))

# plot raster data
plot(SoilP,
     main = "Soil P")
```

Convert raster values to data frame
```{r}
SoilP <- brick(here(ebf_loc,'SoilP_EBFOnly.tif'))
```

```{r}
SoilP_df <- as.data.frame(rasterToPoints(SoilP,spatial = TRUE))
head(SoilP_df)
hist(SoilP_df$SoilP_EBFOnly)
```


```{r}
median_tropics <- median(SoilP_df$SoilP_EBFOnly, na.rm = TRUE)
median_bnf     <- median(bnf.ebf.ex$soil_total_phosphorus_EX, na.rm = TRUE)
median_denit   <- median(denit.ebf$Extracted_soil_total_phosphorus, na.rm = TRUE)

# --- 2. Set axis limits and color palette ---
x_axis_extent <- c(0, 1250)

color_palette <- c(
  "EBFs" = "grey",
  "BNF data" = "red",
  "Denitrification \ndata" = "blue"
)

# --- 3. Create the plot with density curves and vertical dashed lines ---
p <- ggplot() +
  geom_density(
    data = SoilP_df,
    aes(x = SoilP_EBFOnly,
        fill = factor("EBFs", levels = names(color_palette))),
    alpha = 0.35
  ) +
  geom_density(data = bnf.ebf.ex, aes(x = soil_total_phosphorus_EX, fill = "BNF data"), alpha = 0.35) +
  geom_density(data = denit.ebf, aes(x = Extracted_soil_total_phosphorus, fill = "Denitrification \ndata"), alpha = 0.35) +
  
  # Add vertical dashed lines at density peaks (modes)
  geom_vline(xintercept = median_tropics, linetype = "dashed", color = "grey20", linewidth = 0.75) +
  geom_vline(xintercept = median_bnf,     linetype = "dashed", color = color_palette["BNF data"], linewidth = 0.75) +
  geom_vline(xintercept = median_denit,   linetype = "dashed", color = color_palette["Denitrification \ndata"], linewidth = 0.75) +

  # Labels, theme, and scale adjustments
  labs(title = "Within evergreen broadleaf forests", x = expression("Soil P (mg kg"^{-1}*")"), y = " ") +
  theme_minimal() +
  xlim(x_axis_extent) +
  scale_fill_manual(values = color_palette, name = " ") +
  theme(
    legend.position = "none",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.title = element_text(size = 20),
    axis.text = element_text(size = 18),
    plot.title = element_text(size = 22, face = "bold", hjust = 0.5)  # Change font size and make it bold
  )

# Show plot
p
#ggsave(here(figure_loc, "MedianDensity/SoilP_EBFOnly_Density_expanded.png"), plot = p, width = 8, height = 6, dpi = 400)
```









##Checking colinearity of environemntal factors

```{r}
#remove NA values and test for colinearity 
 data[data==0] <- NA
 data
 
 raster_data_noZeros<-data[complete.cases(data),]


# We can visually look for correlations between variables:
heatmap(abs(cor(raster_data_noZeros)), 
        # Compute pearson correlation (note they are absolute values)
        col = rev(heat.colors(6)), 
        Colv = NA, Rowv = NA)
legend("topleft", 
       title = "Absolute Pearson R",
       legend =  round(seq(0,1, length.out = 6),1),
       y.intersp = 0.7, bty = "n",
       fill = rev(heat.colors(6)))
```

```{r}
 #raster_data[raster_data==0] <- NA
 #raster_data
 
# raster_data_noZeros<-raster_data[complete.cases(raster_data),]

# Prepare correlation matrix
corr <- abs(cor(raster_data_noZeros))

# Set layout: 1 row, 2 columns (plot + legend)
layout(matrix(c(1, 2), nrow = 1), widths = c(4, 1))

# Plot heatmap
heatmap(corr, col = rev(heat.colors(6)), Colv = NA, Rowv = NA)

# Plot empty area and draw legend
par(mar = c(5, 1, 4, 4))
plot.new()
legend("center", 
       title = "Abs Pearson R",
       legend = round(seq(0, 1, length.out = 6), 1),
       fill = rev(heat.colors(6)), 
       bty = "n")
```






##Principal Component Analysis 
only using those factors that have an absolute pearson R less than 0.8 
```{r}
data <- read.csv(here(data_loc,'RandPoints_ext.csv'), header= TRUE)
data <- data %>% dplyr::select(-c(NEE, IGBP, GPP, Elevation, Roughness, SoilpH, PPT))

data <- sapply( data, as.numeric, rm.na = TRUE)
data <- as.data.frame(data)

```

```{r}
for (col in 1:ncol(data)) {
    hist(data[,col], main=colnames(data)[col],  xlab=colnames(data)[col])
}
```


Normalizing values 
```{r}


#data$PPT <- (data$PPT) +0.001

#data$PPT <- log(data$PPT)

data$MAT <- (data$MAT) +0.001

data$MAT <- log(data$MAT)

#data$GPP <- (data$GPP) +0.001

#data$GPP <- log(data$GPP)

data$NPP <- (data$NPP) +0.001

data$NPP <- log(data$NPP)

data$Ndep2050 <- (data$Ndep2050) +0.001

data$Ndep2050 <- log(data$Ndep2050)

data$Slope <- (data$Slope) +0.001

data$Slope <- log(data$Slope)

#data$Elevation <- (data$Elevation) +0.001

#data$Elevation <- log(data$Elevation)

#data$Roughness <- (data$Roughness) +0.001

#data$Roughness <- log(data$Roughness)

data$SoilOC <- (data$SoilOC) +0.001

data$SoilOC <- log(data$SoilOC)

data$SoilN <- (data$SoilN) +0.001

data$SoilN <- log(data$SoilN)

data$SoilP <- (data$SoilP) +0.001

data$SoilP <- log(data$SoilP)

#data$SoilpH <- (data$SoilpH) +0.001

#data$SoilpH <- log(data$SoilpH)

data$DtB1 <- (data$DtB1) +0.1

data$DtB1 <- log(data$DtB1)
```

```{r}
for (col in 1:ncol(data)) {
    hist(data[,col], main=colnames(data)[col],  xlab=colnames(data)[col])
}
```

Based on QCBS vegan notes (https://r.qcbs.ca/workshop09/book-en/learning-objectives.html)
FactoMineR documentation

```{r}
# Frist, normalize and then standardize your data (data previously normalized and labels removed)

#soil_var.2 <- decostand(soil_var.2, "standardize")

raster_data <- decostand(data, "standardize")
```

```{r}
# Perform PCA with FactoMineR 
# visualize with factoextra

raster.pca <- FactoMineR::PCA(raster_data)
 
summary(raster.pca)  

  ## unknown scaling for the below plot
```

```{r}
# Control variable colors using their contribution
# Possible values for the argument col.var are :
  # "cos2", "contrib", "coord", "x", "y"
fviz_pca_var(raster.pca, col.var="contrib") +
scale_color_gradient2(low="blue", mid="yellow", 
                      high="orange", midpoint=5)+theme_bw()
```

```{r}
raster.pca.plot <- factoextra::fviz_pca_biplot(raster.pca, # PCA object
                axes = c(1, 2), # graph first two PCs
                geom.ind = "point", # display trees as points
                geom.var = c("arrow", "text"), # add arrows and text to variables
                pointsize = 2, # increase size of tree points
                labelsize = 6, # increase graph label font size
                alpha = 0.5, # fade points/ellipses to read better
                arrowsize = 1,
                #habillage = as.factor(soil_samp$site), 
                # habillage = as.factor(soil_clusters),
                addEllipses = FALSE, # draw ellipses around groups (sites)
                ellipse.level = 0.95,
                #palette = pal3, # match palette with other plots
                col.ind ="grey",
                col.var = "blue", # make arrows/text of var black
                ggrepel = TRUE, # avoid label overlap
                mean.point = FALSE) + # remove site average point
        theme(panel.grid = element_blank(),
        legend.title = element_blank()) +
  geom_text(aes(x= -5, y = 4), label = "", size = 4) +
  theme(text = element_text(size = 17),
        axis.title = element_text(size = 17),
        axis.text = element_text(size = 17),
        title = element_blank()) 

raster.pca.plot

# export plot to output folder

#png("All_PCA_2024-4-09.png", height = 8, width = 12, unit = "in", res = 800)
#print(raster.pca.plot)
#dev.off()
```

```{r}
# obtain correlation coefficients between variables and
# PCs, and run a significance test on correlations

FactoMineR::dimdesc(raster.pca, axes = c(1, 2, 3))
    ## all variables are significantly correlated with PCS
```
```{r}
loadings <- raster.pca$var$coord
as.data.frame(loadings)

file_path <- here(outputs_loc, file = "PCAloadings.csv")
write.csv(loadings, file = file_path, row.names = FALSE)
```


```{r}
## examine variables loadings for PCs
## A reference dashed line is also shown on the barplot. 
## This reference line corresponds to the expected value 
## if the contribution where uniform. Bars above this dashed line are important contributors 

fviz_contrib(raster.pca, choice = "var", axes = 1)
fviz_contrib(raster.pca, choice = "var", axes = 2)
fviz_contrib(raster.pca, choice = "var", axes = 3)  
fviz_contrib(raster.pca, choice = "var", axes = 4) 

# rm(soil.pca.plot, soil.pca)

```

```{r}

# Kaiser Guttman criteria - select PCs that explain the more than
# the average across all PCs

ev <- raster.pca$eig[ ,1] # select eigenvalues for each variable
barplot(ev)
abline(h = mean(ev), col = "red3")
legend("topright", "Average Eigenvalue", col = "red3")

# useful information down to PC3

```

```{r}
#Make a scree plot using base graphics : A scree plot is a graph of the eigenvalues/variances associated with components.
eigenvalues <- raster.pca$eig
barplot(eigenvalues[, 2], names.arg=1:nrow(eigenvalues), 
       main = "Variances",
       xlab = "Principal Components",
       ylab = "Percentage of variances",
       col ="steelblue")
# Add connected line segments to the plot
lines(x = 1:nrow(eigenvalues), eigenvalues[, 2], 
      type="b", pch=19, col = "red")
```

##Reattaching locations and categorical variables to PC scores 
```{r}
pc_scores <- as.data.frame(raster.pca$ind$coord)
```

```{r}
data2 <- read.csv(here(pca_loc, 'RandPoints_Ext_locations.csv'), header= TRUE)
```

```{r}
with_loc <- merge(data2, pc_scores, by = 0)
#write.csv(with_loc, file = "PCScores_noColinear_5-10-24.csv")
```

```{r}
with_loc <- subset(with_loc, Dim.2 < 20, select = Row.names:Dim.5)
```

```{r}
cols <- c("#F72424", "#0000FF", "#A0A0A050")
ordered <- with_loc[order(with_loc$Type, decreasing = TRUE), ]

points <- ggplot(ordered, aes(x = Dim.1, y = Dim.2, color = Type, order = Type)) +
  geom_hline(yintercept = 0, linetype = "solid", linewidth = 1, colour = "black") +  # Bold zero horizontal line
  geom_vline(xintercept = 0, linetype = "solid", linewidth = 1, colour = "black") +
  geom_point() +
  scale_color_manual(values = cols) + 
  theme_bw()+
  xlab("PC1 (41%)") + ylab("PC2 (14%)")+
  theme(
    legend.position = "none",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.title = element_text(size = 16), # Adjust font size for axis titles
    axis.text = element_text(size = 14),   # Adjust font size for axis text
    legend.title = element_text(size = 14), # Legend title font size
    legend.text = element_text(size = 12)   # Legend item font size
  )
   # Bold zero vertical line
points

#ggsave(here(figure_loc, "PC_Scores_PointsOver_NoLegend.png"), plot = points, width = 8, height = 6, dpi = 300)
```

```{r}
# plot
points.pca <- ggplot(ordered, aes(x = Dim.1, y = Dim.2, color = Type, order = Type)) +
  geom_hline(yintercept = 0, linetype = "solid", linewidth = 1, colour = "black") +  # Bold zero horizontal line
  geom_vline(xintercept = 0, linetype = "solid", linewidth = 1, colour = "black") +
  geom_point() +
  scale_color_manual(values = cols) + 
  theme_bw()+
  xlab("PC1 (41%)") + ylab("PC2 (14%)")+
  theme(
    legend.position = "none",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.title = element_text(size = 16), # Adjust font size for axis titles
    axis.text = element_text(size = 14),   # Adjust font size for axis text
    legend.title = element_text(size = 14), # Legend title font size
    legend.text = element_text(size = 12)   # Legend item font size
  )

# Add density curves to y and x axis
#xdens <- 
  #axis_canvas(points.pca, axis = "x") + 
  #geom_density(data = ordered, aes(x = Dim.1, fill = Type, color = Type), alpha = 0.3)

xdens <- 
  axis_canvas(points.pca, axis = "x") + 
  geom_density(data = ordered, 
               aes(x = Dim.1, fill = Type, color = Type), 
               alpha = 0.3) +
  scale_fill_manual(values = cols) +
  scale_color_manual(values = cols)

ydens <-
  axis_canvas(points.pca, axis = "y", coord_flip = TRUE) + 
  geom_density(data = ordered, aes(x = Dim.2, fill = Type, color = Type), alpha = 0.3) +
  coord_flip() +
  scale_fill_manual(values = cols) +
  scale_color_manual(values = cols)

points.pca <- points.pca %>%
  insert_xaxis_grob(xdens, grid::unit(1, "in"), position = "top") %>%
  insert_yaxis_grob(ydens, grid::unit(1, "in"), position = "right") %>%
  ggdraw()
points.pca

#ggsave(here(figure_loc, "PC_Scores_PointsOver_Density.png"), plot = points.pca, width = 8, height = 6, dpi = 400)
```





```{r}
# plot
points.pca <- ggplot(ordered, aes(x = Dim.1, y = Dim.3, color = Type, order = Type)) +
  geom_hline(yintercept = 0, linetype = "solid", linewidth = 1, colour = "black") +  # Bold zero horizontal line
  geom_vline(xintercept = 0, linetype = "solid", linewidth = 1, colour = "black") +
  geom_point() +
  scale_color_manual(values = cols) + 
  theme_bw()+
  xlab("PC1 (41%)") + ylab("PC3 (10.3%)")+
  theme(
    legend.position = "none",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.title = element_text(size = 16), # Adjust font size for axis titles
    axis.text = element_text(size = 14),   # Adjust font size for axis text
    legend.title = element_text(size = 14), # Legend title font size
    legend.text = element_text(size = 12)   # Legend item font size
  )

# Add density curves to y and x axis
#xdens <- 
  #axis_canvas(points.pca, axis = "x") + 
  #geom_density(data = ordered, aes(x = Dim.1, fill = Type, color = Type), alpha = 0.3)

xdens <- 
  axis_canvas(points.pca, axis = "x") + 
  geom_density(data = ordered, 
               aes(x = Dim.1, fill = Type, color = Type), 
               alpha = 0.3) +
  scale_fill_manual(values = cols) +
  scale_color_manual(values = cols)

ydens <-
  axis_canvas(points.pca, axis = "y", coord_flip = TRUE) + 
  geom_density(data = ordered, aes(x = Dim.3, fill = Type, color = Type), alpha = 0.3) +
  coord_flip() +
  scale_fill_manual(values = cols) +
  scale_color_manual(values = cols)

points.pca <- points.pca %>%
  insert_xaxis_grob(xdens, grid::unit(1, "in"), position = "top") %>%
  insert_yaxis_grob(ydens, grid::unit(1, "in"), position = "right") %>%
  ggdraw()
points.pca

ggsave(here(figure_loc, "PC_Scores_PointsOver_Density_OnevsThree.png"), plot = points.pca, width = 8, height = 6, dpi = 400)
```






##Generalized Additive Model (GAM)
```{r}
df <- read.csv(here(pca_loc,'RandPoints_Ext.csv'), header= TRUE)
df$Slope <- as.numeric(df$Slope)
df <- df %>% dplyr::select(-c(NEE, IGBP, GPP, Elevation, Roughness, SoilpH, PPT))
```



```{r}
#read in data and standardize variables 
numeric_columns <- sapply(df, is.numeric)
df[numeric_columns] <- scale(df[numeric_columns])
```

```{r}
#attach them to the presence/absence classification
gam_data <- read.csv(here(pca_loc,"RandPoints_Ext_all_GAM_loc.csv"), header = TRUE)
gam_data <- merge(gam_data, df, by = 0)
gam_data <- gam_data %>% dplyr::select(-c(Row.names))
```

```{r}
#remove NA rows 
gam_data <- na.omit(gam_data)
```

```{r}
hist(gam_data$PresAbs)
```

```{r}
#try to downsample data 
##Remove pres abs data rows first 
# Specify the column and value to remove
column_name <- "PresAbs"
value_to_remove <- "1"

# Separate the rows to be removed
rows_to_remove <- gam_data[gam_data[[column_name]] == value_to_remove, ]

# Data frame without the specified rows
df_remaining <- gam_data[gam_data[[column_name]] != value_to_remove, ]
```

```{r}
#actual downsample
if (nrow(df_remaining) >= 10000) {
  # Randomly sample 10,000 rows
  sampled_df <- df_remaining[sample(nrow(df_remaining), 10000), ]
} else {
  stop("Not enough rows with the specified value to sample 10,000 rows.")
}
```

```{r}
#reattach 
gam_data <- rbind(sampled_df, rows_to_remove)
```

Doing the GAM with 10,000 samples and considering potentially interesting variables: 
```{r}
gam_model<-gam(PresAbs~ s(SoilN) + s(SoilOC) + s(NPP+MAT) + s(Slope) + s(SoilP) + s(AET) + s(SoilMoisture) + s(Seasonality), family = binomial, data=gam_data, select = TRUE, method = "REML")
#Then when you want to predict, you just:
Gam_pred <- predict.gam(gam_model, newdata=gam_data, type="response")
Gam_pred <- as.data.frame(Gam_pred)
```


```{r}
summary(gam_model)$p.table
```


```{r}
hist(Gam_pred$Gam_pred)
```

```{r}
summary(gam_model)
basic_summary <- summary(gam_model)
#already cross validated 
```

```{r}
gam.check(gam_model)
```




```{r}
#Cross-validation
# Perform k-fold cross-validation
gam_data$PresAbs <- as.factor(gam_data$PresAbs)
train_control <- trainControl(method = "cv",
                              number = 10)
cv_model <- train(PresAbs ~ SoilN+SoilOC+NPP+MAT+Slope+SoilP+DtB1+AET+SoilMoisture+Seasonality, data = gam_data, method = "gam", trControl = train_control)
print(cv_model)
```
Accuracy indicates that the model did well to predict negative and positive values. Kapps indicates there is fair agreement, I think meaning that our model is 25% better than what would align only by chance. 


Attaching the GAM results to locations
```{r}
gam_data <- merge(gam_data, Gam_pred, by = 0)
gam_data <- gam_data %>% dplyr::select(-c(Row.names))
```

```{r}
file_path <- here(data_loc, "Gam_10000_output.csv")
write.csv(gam_data, file = file_path, row.names = FALSE)
```



##TRENDY-N Models 
```{r}
modeldata <- here("Data", "TRENDYN")
CP_NMin <- rast(here(loc, "CABLE-POP/fNnetmin_annually.nc"))
CLASSIC_NMin <- rast(here(loc,"CLASSIC/fNnetmin_annually.nc"))
CLM5_NMin <- rast(here(loc,"CLM5.0/fNnetmin_annually.nc"))
DLEM_NMin <- rast(here(loc,"DLEM/fNnetmin_annually.nc"))
#ISAM_NMin <- rast(here(loc,"ISAM/fNnetmin_annually.nc"))
JSBACH_NMin <- rast(here(loc,"JSBACH/fNnetmin_annually.nc"))
JULES_ES_NMin <- rast(here(loc,"JULES-ES/fNnetmin_annually.nc"))
LPJ_GUESS_NMin <- rast(here(loc,"LPJ-GUESS/fNnetmin_annually.nc"))
LPX_Bern_NMin <- rast(here(loc,"LPX-Bern/fNnetmin_annually.nc"))
OCNv2_NMin <- rast(here(loc,"OCNv2/fNnetmin_annually.nc"))
ORCHIDEEv3_NMin <- rast(here(loc,"ORCHIDEEv3/fNnetmin_annually.nc"))
```

BNF - Run this block of code for each individual model, ensuring to change the name of the output file! 
```{r}
Classic_bnf_nc <- ncdf4::nc_open(here(modeldata, "CLASSIC/fBNF_annually.nc"))

data_var <- ncvar_get(Classic_bnf_nc, "fBNF")
time_var <- ncvar_get(Classic_bnf_nc, "time")

lat <- ncvar_get(Classic_bnf_nc, "latitude")
lon <- ncvar_get(Classic_bnf_nc, "longitude")

nlat <- length(lat)
nlon <- length(lon)
ntime <- length(time_var)

selected_time_steps <- c(150:170)

rasters <- list()

for (i in selected_time_steps) {
  data_slice <- data_var[,,i]
  r <- raster(t(data_slice), xmn=min(lon), xmx=max(lon), ymn=min(lat), ymx=max(lat), crs=CRS("+proj=longlat +datum=WGS84"))
  rasters[[length(rasters) + 1]] <- r
}

raster_brick <- brick(rasters)

average_raster <- calc(raster_brick, mean)


##CONVERTS UNITS

average_raster <- flip(average_raster, direction = 'y')
average_raster <- terra::rotate(average_raster)
average_raster_units <- average_raster*(3.1104e+10)
plot(average_raster_units)
#average_raster

values(average_raster_units)

# Define the output file path
output_file <- here(bnfrasters_loc,"Classic_bnf_Avg.tif")

# Write the adjusted average raster to a GeoTIFF file
writeRaster(average_raster_units, output_file, format = "GTiff", overwrite = TRUE)

```

```{r}
# List all raster files (e.g., .tif) in the folder
raster_files <- list.files(BNF, pattern = "\\.tif$", full.names = TRUE)

# Load and stack the rasters
raster_stack <- rast(raster_files)  # This creates a SpatRaster with multiple layers
```

```{r}
# Calculate the mean across the raster stack, removing NA values
mean_raster <- app(spat_raster_stack, fun = function(x) mean(x, na.rm = TRUE))
output_file <- here("Outputs", "Mean_TRENDYN_BNF.tif")
writeRaster(mean_raster, output_file, filetype = "GTiff", overwrite = TRUE)


# Calculate the standard deviation across the raster stack, removing NA values
sd_raster <- app(spat_raster_stack, fun = function(x) sd(x, na.rm = TRUE))
output_file <- here("Outputs","SD_TRENDYN_BNF.tif")
writeRaster(sd_raster, output_file, overwrite = TRUE)

# Calculate the coefficient of variation (CV)
cv_raster <- sd_raster / mean_raster
output_file <- here("Outputs","CV_TRENDYN_BNF.tif")
writeRaster(cv_raster, output_file, overwrite = TRUE)

# Plot the CV raster
plot(sd_raster)
```


N loss - Run this block of code for each individual model, ensuring to change the name of the output file!
```{r}
#averages_loc_nloss <- here("Outputs", "AvgRasters", "Nloss")

Classic_nloss_nc <- ncdf4::nc_open(here(loc, "ORCHIDEEv3/fNloss_annually.nc"))

data_var <- ncvar_get(Classic_nloss_nc, "fNloss")
time_var <- ncvar_get(Classic_nloss_nc, "time")

lat <- ncvar_get(Classic_nloss_nc, "latitude")
lon <- ncvar_get(Classic_nloss_nc, "longitude")

nlat <- length(lat)
nlon <- length(lon)
ntime <- length(time_var)

selected_time_steps <- c(150:170)

rasters <- list()

for (i in selected_time_steps) {
  data_slice <- data_var[,,i]
  r <- raster(t(data_slice), xmn=min(lon), xmx=max(lon), ymn=min(lat), ymx=max(lat), crs=CRS("+proj=longlat +datum=WGS84"))
  rasters[[length(rasters) + 1]] <- r
}

raster_brick <- brick(rasters)

average_raster <- calc(raster_brick, mean)




average_raster <- flip(average_raster, direction = 'y')
average_raster <- terra::rotate(average_raster)
average_raster_units <- average_raster*(3.1104e+10)
plot(average_raster_units)
#average_raster

values(average_raster_units)

# Define the output file path
output_file <- here(averages_loc_nloss,"ORCHIDEEv3_Nloss_Avg.tif")

# Write the adjusted average raster to a GeoTIFF file
writeRaster(average_raster_units, output_file, format = "GTiff", overwrite = TRUE)

```

```{r}
# Calculate the mean across the raster stack, removing NA values
mean_raster <- app(spat_raster_stack, fun = function(x) mean(x, na.rm = TRUE))
output_file <- "Mean_TRENDYN_Nloss.tif"
writeRaster(mean_raster, output_file, overwrite = TRUE)

# Calculate the standard deviation across the raster stack, removing NA values
sd_raster <- app(spat_raster_stack, fun = function(x) sd(x, na.rm = TRUE))
output_file <- "SD_TRENDYN_Nloss.tif"
writeRaster(sd_raster, output_file, overwrite = TRUE)

# Calculate the coefficient of variation (CV)
cv_raster <- sd_raster / mean_raster
output_file <- "CV_TRENDYN_Nloss.tif"
writeRaster(cv_raster, output_file, overwrite = TRUE)

# Plot the sd raster
plot(sd_raster)
```





##Plotting density curves with TRENDY data
```{r}
bnf <- read.csv(here(data_loc,'BNF_TRENDYN.csv'), header= TRUE)
bnf_no_NA <- bnf %>% drop_na(Variance)

denit <- read.csv(here(data_loc,'Nloss_TRENDYN.csv'), header= TRUE)
denit_no_NA <- denit %>% drop_na(Variance)
```

```{r}
#Natural
#model_bnf <- raster(x = "CV_TRENDYN_BNF.tif")
mean_bnf_nat <- brick(here(modelsNat_loc,"BNF_M_Nat.tif"))
mean_bnf_nat_df <- as.data.frame(rasterToPoints(mean_bnf_nat,spatial = TRUE))
```

```{r}
x_axis_extent <- c(0,7.5)

# Specify the colors for each dataset
color_palette <- c("BNF - Tropics" = "grey", "BNF - Sampled points" = "red")

# Create density plots using ggplot2 with specified x-axis extent, colors, and legend
p <- ggplot() +
  geom_density(data = mean_bnf_nat_df, aes(x = BNF_M_Nat, fill = "BNF - Tropics"), alpha = 0.5) +
  geom_density(data = bnf_no_NA, aes(x = Mean, fill = "BNF - Sampled points"), alpha = 0.5) +
  labs(
  x = expression("Mean BNF rate (kg N yr"^{-1}*")"),
  y = "Density") +
  theme_minimal() +
  xlim(x_axis_extent) +
  scale_fill_manual(values = color_palette, name = " ") +  # Set legend name and colors
   theme(
    legend.position = c(0.83, 0.86), # Adjust legend position
    axis.title = element_text(size = 16), # Adjust font size for axis titles
    axis.text = element_text(size = 14)   # Adjust font size for axis text
  )
p
#ggsave("Mean_BNF_Clean.png", plot = p, width = 8, height = 6, dpi = 300)
```

```{r}
# --- 1. Calculate modes (peak positions) for each group ---
# Remove NAs if needed
median_nat <- median(mean_bnf_nat_df$BNF_M_Nat, na.rm = TRUE)
median_bnf     <- median(bnf_no_NA$Mean, na.rm = TRUE)

# --- 2. Set axis limits and color palette ---
x_axis_extent <- c(0, 4.5)

color_palette <- c(
  "BNF - Tropics" = "grey",
  "BNF - Sampled points" = "red"
)

# --- 3. Create the plot with density curves and vertical dashed lines ---
p <- ggplot() +
  geom_density(data = mean_bnf_nat_df, aes(x = BNF_M_Nat, color = "BNF - Tropics"), linewidth = 1.5) +
  geom_density(data = bnf_no_NA, aes(x = Mean, color = "BNF - Sampled points"), linewidth = 1.5) +
  
  # Add vertical dashed lines at density peaks (modes)
  geom_vline(xintercept = median_nat, linetype = "dashed", color = color_palette["BNF - Tropics"], linewidth = 0.75) +
  geom_vline(xintercept = median_bnf,     linetype = "dashed", color = color_palette["BNF - Sampled points"], linewidth = 0.75) +

  # Labels, theme, and scale adjustments
  labs(
    x = expression("g N m"^{-2}*" yr"^{-1}*""),
    #x = " ",
    y = "Density") +
  theme_minimal() +
  xlim(x_axis_extent) +
  ylim(0, 0.675) +  # Lock y-axis extent
  scale_color_manual(values = color_palette, name = " ") +
  theme(
    legend.position = "none",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.text = element_text(size = 14),            # Increase legend text size
    legend.key.size = unit(1.0, "cm"),                # Make legend boxes larger
    axis.title = element_text(size = 25),
    axis.line = element_line(color = "black"),
    axis.text = element_text(size = 25),
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5)
  )

# Show plot
p
#ggsave(here(figure_loc, "MeanBNFRate_Density_Natural_empty.png"), plot = p, width = 8, height = 6, dpi = 400)
```















